### 2023-10-25 post_prop.tex

\documentclass[10pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=.5in, footskip=0.5in]{geometry}
\usepackage{graphicx}

\begin{document}
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/dimensionalities_post.pdf}
\end{center}
\end{figure}
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/endpoints_post.pdf}
\end{center}
\end{figure}

\newpage
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/dimensionalities_post_log.pdf}
\end{center}
\end{figure}
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/endpoints_post_log.pdf}
\end{center}
\end{figure}

\newpage
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/dimensionalities_prop.pdf}
\end{center}
\end{figure}
\begin{figure}
\begin{center}
    \includegraphics{notebooks/temperature/figures/endpoints_prop.pdf}
\end{center}
\end{figure}

\end{document}